---
title: 'Mini Project #3'
author: "Jane Choi"
date: "10/8/2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("/Users/janechoi/Desktop/IU/DATA/miniproject")
library(gapminder)
library(tidyverse)
library(rio)
library(data.table)
library(GGally)
library(mgcv)
library(modelr)
library(broom)
library(arm)
library(plotly)


cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

life.expectancy =fread('life_expectancy_years.csv',header = T)
population<-fread('population_total.csv',header = T)
gdp<-fread('income_per_person_gdppercapita_ppp_inflation_adjusted.csv',header = T)
gapminder <- gapminder


#unique(population$country) #193
#unique(gapminder$country) #142
#unique(gdp$country) #193
#unique(life.expectancy$country) #187

gapminder.new<- gapminder[,c(1,2)]

gapminder.new<- unique(gapminder.new)

new.gdp <- merge(gdp, gapminder.new, by='country')  
new.life.expectancy <- merge(life.expectancy, gapminder.new, by='country') 
new.population <- merge(population, gapminder.new, by='country') 



life.expectancy.long = gather(new.life.expectancy, key = life.expectancy.date , value = life.expectancy.number, 2:220)
gdp.long <- gather(new.gdp, key= gdp.date, value= gdp.number , 2:242)

population.long <- gather(new.population, key=population.date, value=population.number, 2:302)


gdp.long$key <- paste(gdp.long$country , by='-',gdp.long$gdp.date)
population.long$key <- paste(population.long$country , by='-',population.long$population.date)
life.expectancy.long$key <- paste(life.expectancy.long$country , by='-',life.expectancy.long$life.expectancy.date)



data<- full_join(gdp.long, life.expectancy.long, by='key')
data<- full_join(data, population.long, by='key')

data<-data[,-c(6:8,10:12)]

colnames(data) <- c("country" , "continent","date","gdp.number", "key" ,"life.expectancy.number", "population.number")


data<- data[,c(5,3,1,2,4,6,7)]

#head(data)

colSums(is.na(data))

data <- data[is.na(data$life.expectancy.number) == F,]
colSums(is.na(data))




#write.csv(data, file='data.csv',row.names=F)
#data.new<-read.csv('data.csv')

data.ww2 <- data %>%
  filter(data$date > 1939)



```




Making date.factor variable :
```{r}
data.ww2$date.factor <- 0

i=0

for ( i in 1:10586){
  if (data.ww2[i,2] <= 1949 ) {
    data.ww2[i,8] <- '1940s'}
  
  else if (data.ww2[i,2] <= 1959){
     data.ww2[i,8] <- '1950s'
  }
   else if (data.ww2[i,2] <= 1969){
     data.ww2[i,8] <- '1960s'
   }
   else if (data.ww2[i,2] <= 1979){
     data.ww2[i,8] <- '1970s'
   }
   else if (data.ww2[i,2] <= 1989){
     data.ww2[i,8] <- '1980s'
   }
   else if (data.ww2[i,2] <= 1999){
     data.ww2[i,8] <- '1990s'
   }
   else if (data.ww2[i,2] <= 2009){
     data.ww2[i,8] <- '2000s'
   }
   else{ data.ww2[i,8] <- '2010s'
  }
}



data.ww2$date<-as.factor(data.ww2$date)

data.ww2$date.factor<-as.factor(data.ww2$date.factor)

str(data.ww2)

```





Seeing  plots by ggpairs:
```{r}


ggpairs(data.ww2[,c(5:6,8)])


```


Seeing  plots by plotly:
```{r}
#plot_ly(data.ww2, x= ~log(gdp.number) ,y=~life.expectancy.number , z=~date, type= 'scatter3d' , marker= list(size=1.5))
```





##1. Method: Modeling without continent information:

###(1-1) model with linear:

```{r}
data.ww2.lm = lm(life.expectancy.number ~ log10(gdp.number) + date, data = data.ww2 )
data.ww2.lm.df = augment(data.ww2.lm)

colnames(data.ww2.lm.df) = c("life.expectancy.number" , "gdp.number" ,  "date"  ,   ".fitted"  , ".se.fit" ,  ".resid" , ".hat" ,   ".sigma" ,".cooksd"  ,   ".std.resid"  )

ggplot(data.ww2.lm.df, aes(x = gdp.number, y = .resid)) + 
  geom_smooth(se=F) +geom_point(alpha=0.1) +
  ggtitle('Residual Plot for  Linear Model ') +
  labs(x='log10 of gdp')



ggplot(data.ww2.lm.df, aes(x = .fitted, y = abs(.resid))) + 
  geom_smooth(se=F) +geom_point(alpha=0.1)+
  ggtitle('Residual & Fitted Plot for Linear Model ') +
  labs(x='log10 of gdp')

```


-not good residual plot


Make predictions: 
```{r}


ww2.lm.grid <-expand.grid(gdp.number = seq(5.5, 11.6, 0.1), date = seq(1940,2018,1))

ww2.lm.grid$date<-as.factor(ww2.lm.grid$date)


ww2.lm.grid.predict = predict(data.ww2.lm, newdata = ww2.lm.grid)
ww2.lm.grid.predict = data.frame(ww2.lm.grid, life.expectancy = as.vector(ww2.lm.grid.predict))


#ggplot(ww2.lm.grid.predict, aes(x = gdp.number, y = life.expectancy)) +  geom_line() +facet_wrap(~date)



ggplot(ww2.lm.grid.predict, aes(x = gdp.number, y = life.expectancy, group = date, color = date)) + geom_line() +
  ggtitle('Predicted Life Expectancy per GDP after WW2' ) + 
labs(x= 'Log 10 of GDP per capita' , y= 'Life Expectancy', subtitle = 'Using linear model with seperate dates')

```


-looks different in every date / date has effect! 





###(1-2) model with liner : using date.factors


```{r}
data.ww2.lm.f = lm(life.expectancy.number ~ log10(gdp.number) + date.factor, data = data.ww2)
data.ww2.lm.f.df = augment(data.ww2.lm.f)

colnames(data.ww2.lm.f.df) = c("life.expectancy.number" , "gdp.number" ,  "date"  ,   ".fitted"  , ".se.fit" ,  ".resid" , ".hat" ,   ".sigma" ,".cooksd"  ,   ".std.resid"  )

ggplot(data.ww2.lm.f.df, aes(x = gdp.number, y = .resid)) + 
  geom_smooth(se=F) +geom_point(alpha=0.1) +
  ggtitle('Residual Plot for  Linear Model ') +
  labs(x='log10 of gdp')



ggplot(data.ww2.lm.df, aes(x = .fitted, y = abs(.resid))) + 
  geom_smooth(se=F) +geom_point(alpha=0.1)+
  ggtitle('Residual & Fitted Plot for Linear Model ') +
  labs(x='log10 of gdp')

```




Make predictions: 
```{r}


ww2.lm.f.grid <-expand.grid(gdp.number = seq(5.5, 11.6, 0.1), date.factor = c('1940s','1950s','1960s','1970s','1980s','1990s','2010s'))

ww2.lm.f.grid$date.factor<-as.factor(ww2.lm.f.grid$date.factor)


ww2.lm.f.grid.predict = predict(data.ww2.lm.f, newdata = ww2.lm.f.grid)
ww2.lm.f.grid.predict = data.frame(ww2.lm.f.grid, life.expectancy = as.vector(ww2.lm.f.grid.predict))



#ggplot(ww2.lm.f.grid.predict, aes(x = gdp.number, y = life.expectancy)) +  geom_line() +facet_wrap(~date.factor)

ggplot(ww2.lm.f.grid.predict, aes(x = gdp.number, y = life.expectancy, group = date.factor, color = date.factor)) + geom_line()+
  ggtitle('Predicted Life Expectancy per GDP after WW2' ) + 
  labs(x= 'Log 10 of GDP per capita' , y= 'Life Expectancy', subtitle = 'Using linear model with dates together')+
  scale_color_manual(values=cb_palette)

```



Check the R-squared value for each model: 

```{r}
summary(data.ww2.lm) #Adjusted R-squared:  0.7931 
summary(data.ww2.lm.f) #Adjusted R-squared:  0.7877 
```









##2. Method: Modeling with different continent information:

###(2-0) not using date when modeling with linear: 

Write a function to fit a LM for each group:
```{r}
continent.lm.no.date.f = function(data){
  lm(life.expectancy.number ~ log10(gdp.number), data = data)
}
```

Make a nested data frame:

```{r}

continent.lm.m.no.date.f = data.ww2 %>%
  group_by(continent) %>%
  nest()


continent.lm.m.no.date.f.m = map(continent.lm.m.no.date.f$data, continent.lm.no.date.f)

continent.lm.m.no.date.f = mutate(continent.lm.m.no.date.f, model = continent.lm.m.no.date.f.m)

continent.lm.m.no.date.f = mutate(continent.lm.m.no.date.f, .fitted = map2(data, model, add_predictions))

continent.lm.m.no.date.f = mutate(continent.lm.m.no.date.f, .resid = map2(data, model, add_residuals))

continent.lm.m.no.date.f.fitted = unnest(continent.lm.m.no.date.f, .fitted,.resid)



continent.lm.m.no.date.f.fitted<-continent.lm.m.no.date.f.fitted[,-c(10:16)]

```



ploting the  residuals :
```{r}

ggplot(continent.lm.m.no.date.f.fitted , aes(x=(gdp.number),y= resid))+ 
  geom_point()+ geom_smooth(method='loess',se=F)+
  scale_x_log10() +
  ggtitle("Residual Plot for Gdp per Capita and Life Expectancy")+
  labs(x='log 10 of gdp per capita')


ggplot(continent.lm.m.no.date.f.fitted , aes(x=(gdp.number),y= abs(resid))) + 
  geom_point()+ geom_smooth(method='loess',se=F)+
  scale_x_log10() +
  ggtitle("Absolute Residual Plot for Gdp per Capita and Life Expectancy")+
  labs(x='log 10 of gdp per capita')

```




Make model visuable: 

```{r}

continent.lm.m.no.date.f.fitted.df<-as.data.frame(continent.lm.m.no.date.f.fitted)


ggplot(continent.lm.m.no.date.f.fitted.df, aes(x = gdp.number, y = pred, group=continent , color= continent)) +
  geom_line()+scale_x_log10() +
  ggtitle('Predicted Life Expectancy per GDP after WW2' ) + 
  labs(x= 'Log 10 GDP per capita' , y= 'Life Expectancy', subtitle = 'Using linear model by continents')+
  scale_color_manual(values=cb_palette)


```

###(2-1) model with linear:


Write a function to fit a LM for each group:
```{r}
continent.lm = function(data){
  lm(life.expectancy.number ~ log10(gdp.number) + date, data = data)
}
```

Make a nested data frame:

```{r}

continent.lm.m = data.ww2 %>%
  group_by(continent) %>%
  nest()


continent.lm.m.m = map(continent.lm.m$data, continent.lm)

continent.lm.m = mutate(continent.lm.m, model = continent.lm.m.m)
continent.lm.m = mutate(continent.lm.m, .fitted = map2(data, model, add_predictions))
continent.lm.m = mutate(continent.lm.m, .resid = map2(data, model, add_residuals))
continent.lm.m.fitted = unnest(continent.lm.m, .fitted,.resid)

continent.lm.m.fitted<-continent.lm.m.fitted[,-c(10:16)]



```



ploting the  residuals :
```{r}

ggplot(continent.lm.m.fitted , aes(x=(gdp.number),y= resid))+ 
  geom_point()+ geom_smooth(method='loess',se=F)+
  scale_x_log10() +
  ggtitle("Residual Plot for Gdp per Capita and Life Expectancy")+
  labs(x='log 10 of gdp per capita')


ggplot(continent.lm.m.fitted , aes(x=(gdp.number),y= abs(resid))) + 
  geom_point()+ geom_smooth(method='loess',se=F)+
  scale_x_log10() +
  ggtitle("Absolute Residual Plot for Gdp per Capita and Life Expectancy")+
  labs(x='log 10 of gdp per capita')

```


Make model visuable: 

```{r}

continent.lm.m.fitted.df<-as.data.frame(continent.lm.m.fitted)


ggplot(continent.lm.m.fitted.df, aes(x = gdp.number, y = pred, group=date , color= date)) +
  geom_line()+scale_x_log10() +
  facet_grid(~continent)+
  ggtitle('Predicted Life Expectancy per GDP after WW2' ) + 
  labs(x= 'Log 10 GDP per capita' , y= 'Life Expectancy', subtitle = 'Using linear model faceting by continents')





```



###(1-2) model with liner : using date.factors

Write a function to fit a LM for each group:
```{r}
continent.lm.f = function(data){
  lm(life.expectancy.number ~ log10(gdp.number) + date.factor, data = data)
}
```

Make a nested data frame:

```{r}
continent.lm.m.f = data.ww2 %>%
  group_by(continent) %>%
  nest()

continent.lm.m.f.m = map(continent.lm.m.f$data, continent.lm.f)

continent.lm.m.f = mutate(continent.lm.m.f, model = continent.lm.m.f.m)
continent.lm.m.f = mutate(continent.lm.m.f, .fitted = map2(data, model, add_predictions))
continent.lm.m.f = mutate(continent.lm.m.f, .resid = map2(data, model, add_residuals))
continent.lm.m.f.fitted = unnest(continent.lm.m.f, .fitted,.resid)

continent.lm.m.f.fitted<-continent.lm.m.f.fitted[,-c(10:16)]



```



plotting the  residuals :
```{r}

ggplot(continent.lm.m.f.fitted , aes(x=(gdp.number),y= resid))+ 
  geom_point()+ geom_smooth(method='loess',se=F)+
  scale_x_log10() +
  ggtitle("Residual Plot for Gdp per Capita and Life Expectancy")+
  labs(x='log 10 of gdp per capita')


ggplot(continent.lm.m.f.fitted , aes(x=(gdp.number),y= abs(resid))) + 
  geom_point()+ geom_smooth(method='loess',se=F)+
  scale_x_log10() +
  ggtitle("Absolute Residual Plot for Gdp per Capita and Life Expectancy")+
  labs(x='log 10 of gdp per capita')

```




Visual graph:

```{r}

ggplot(continent.lm.m.f.fitted, aes(x = gdp.number, y = pred, group = date.factor, color = date.factor)) +
  geom_line()+scale_x_log10()+
  facet_grid(~continent)+
  ggtitle('Predicted Life Expectancy per GDP after WW2' ) + 
  labs(x= 'Log 10 of GDP per capita' , y= 'Life Expectancy', subtitle = 'Using linear model considering continents and seperate dates', caption = 'Faceted by Continent')+scale_color_manual(values=cb_palette)


ggplot(continent.lm.m.f.fitted, aes(x = gdp.number, y = pred, group = continent, color = continent)) +
  geom_line()+scale_x_log10()+
  facet_grid(~date.factor)+
  ggtitle('Predicted Life Expectancy per GDP after WW2' ) + 
  labs(x= 'Log 10 of GDP per capita' , y= 'Life Expectancy', subtitle = 'Using linear model considering continents with dates together', caption = 'Faceted by Date') +scale_color_manual(values=cb_palette)

```


-As the date increases the GDP as well increase! 
-The GDP differs among continents too 
-And the life expectancy increases as well as the gdp increases 


